---
title: "Control structures, solutions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Control structures, solutions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## If-else statement

1. load in the Orange dataset
2. Write a for loop to iterate over the trees (Tree 1 to 5) of the dataset and print the circumference for each
3. Write an if else statement that classifies a tree based on its circumference: 
    <= 65.5 -- small
    >65.5 & <= 161.5 -- medium
    >161.5 large
    try with x=45, x=124 and x=179
4. Do the same with ifelse(), using the categories
    <=  161.5 -- small
    >161.5 large
5. additional: write a for cycle to iterate around the rows of the dataset and assins the category to each row. Create an empty column first to store the categories.     

```{r}
data("Orange")

Hmisc::describe(Orange)

for (i in 1:5) print(Orange$circumference[Orange$Tree==i])

x=45

if (x <= 65.5){
  print("small")
} else if (x>65.5 & x<=161.5){
  print("medium")
} else {
  print("large")
}

Orange$cc <- ifelse(Orange$circumference<=161.5, "small",  "large")

Orange$cc <- NA

for (i in 1:nrow(Orange)){
  print(i)
if (Orange$circumference[i] <= 65.5){
  Orange$cc[i] <- "small"
} else if (Orange$circumference[i]>65.5 & Orange$circumference[8]<=161.5){
  Orange$cc[i] <- "medium"
} else {
  Orange$cc[i] <- "large"
}
}

Orange$cc <- ifelse(Orange$circumference<=65.5, "small", 
                    ifelse(Orange$circumference<=161.5, "medium", "large"))


```


## For, while loop

1. for each tree (1 to 5), print the ages. 
2. for each tree, print the age, in which they were small. 
3. load esoph dataset. 
4. Using for loop(s), summarize ncontrols by agegp and alcgp, skipping the category alcgp=="120+" & agegp=="25-34".
5. Using a for loop simulate the flip a coin twenty times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector.

```{r}

Orange[Orange$Tree==1,"age"]

for (i in 1:5){
  cat("Tree: ", i, " Ages: ", paste0(Orange[Orange$Tree==i,"age"], collapse = ", "), "\n")
}

for (i in 1:5){
  cat("Tree: ", i, " was small at age: ", paste0(Orange[Orange$Tree==i & Orange$cc=="small","age"], collapse = ", "), "\n")
}

# 0     1     2     3     4
# 1     0     1     2     3
# 2     1     0     1     2
# 3     2     1     0     1
# 4     3     2     1     0
data("esoph")
for (agegrp in levels(esoph$agegp)){
  print(sum(esoph$ncontrols[esoph$agegp==agegrp]))
}

for (agegrp in levels(esoph$agegp)){
  for (alcgrp in levels(esoph$alcgp)){
   #print(agegrp)
    #print(alcgrp)
  cat("Age group: ", agegrp, ", alcohol consumption group: ", alcgrp, " number of controls: ", sum(esoph$ncontrols[esoph$alcgp==alcgrp & esoph$agegp==agegrp]), "\n")
  }
}
for (agegrp in levels(esoph$agegp)){
  for (alcgrp in levels(esoph$alcgp)){
    if (agegrp=="25-34" & alcgrp=="120+"){
      #next
    } else {
  cat("Age group: ", agegrp, ", alcohol consumption group: ", alcgrp, " number of controls: ", sum(esoph$ncontrols[esoph$alcgp==alcgrp & esoph$agegp==agegrp]), "\n")}
  }
}

results <- sample(c(0,1), 20, replace = T)

```


## apply, lapply, etc...

1. define the following dataset:

```{r}
dataset1 <- cbind(observationA = 16:8, observationB = c(20:19, 6:12))
```

2. Calculate row and column means with apply 
3. Use apply to multiply the whole table by 3. 

```{r}

rowm <- apply(dataset1, 1, mean, na.rm=T)
colm <- apply(dataset1, 2, mean, na.rm=T)


```


```{r}
#two ways

multiply <- function(x, number){
  x <- x*number
  x
}
res <- apply(dataset1,2, multiply, number=3)

res <- apply(dataset1,2, function(x) x*3)

```

4. create a list with 3 list elements. Print the length of each element. 

```{r}
lll <- list(list(1,3,5,6,1), list(5,2,1), list(6,2,5,7,3,1,9))
```


```{r}
lapply(lll, length)
```

