---
title: "Control structures, exercises"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Control structures, exercises}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## If-else statement

1. load in the Orange dataset
2. Write a for loop to iterate over the trees (Tree 1 to 5) of the dataset and print the circumference for each
```{r}



```

3. Write an if else statement that classifies a tree based on its circumference: 
    <= 65.5 -- small
    >65.5 & <= 161.5 -- medium
    >161.5 large
    try with x=45, x=124 and x=179
4. Do the same with ifelse(), using the categories
    <=  161.5 -- small
    >161.5 large
5. additional: write a for cycle to iterate around the rows of the dataset and assigns the category to each row. Create an empty column first to store the categories.     

```{r}


```


## For  loop

1. for each tree (1 to 5), print the ages. 
2. for each tree, print the age, in which they were small. 

```{r}

```

3. load esoph dataset. 
4. Using for loop(s), table ncontrols by agegp and alcgp, skipping the category alcgp=="120+" & agegp=="25-34", using an if else statement.
5. Using a for loop simulate the flip a coin twenty times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector that you predefine. (sample function)



```{r}


```


## apply, lapply, etc...

1. define the following dataset:

```{r}
dataset1 <- cbind(observationA = 16:8, observationB = c(20:19, 6:12))
```

2. Calculate row and column means with apply 
3. Use apply to find the min and the max of each row and column. 
4. Do the same, but use a for cylce. 

```{r}

```


```{r}

```

4. create a list with 3 list elements. Print the length of each element. 


```{r}



```


6. Multiply each list by 2, find each list's minimum and maximum value. 

```{r}


```

5. Do the same, but use a for cycle on the list. 
6. add names to each listelements. The names should be big, if the element is larger than the medium of the list and small otherwise. The medium should be medium. 


