---
title: "Control structures, exercises"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Control structures, exercises}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## If-else statement

1. load in the Orange dataset
2. Create a new, empty column
3. Write a for loop to iterate over the rows of the dataset and print the circumference
3. Write an if else statement that classifies trees based on their circumference: 
    <= 65.5 -- small
    >65.5 & <= 161.5 -- medium
    >161.5 large
4. Do the same with ifelse()

```{r}
data("Orange")

Hmisc::describe(Orange)


for (i in 1:nrow(Orange)) {
  if (Orange[i, "circumference"]<= 65.5){
    Orange[i, "size"] <- "small"} else if (Orange[i, "circumference"]>161.5) {
       Orange[i, "size"] <- "large"
      
    } else {    Orange[i, "size"] <- "medium"}
}

Orange$size <- ifelse(Orange$circumference<=65.5, "small", ifelse(Orange$circumference>161.5, "large", "medium"))

```


## For  loop

1. for each tree (1 to 5), print the ages. 
2. for each tree, print the age, in which they were small. 
3. load esoph dataset. 
4. Using for loop(s), table ncontrols by agegp and alcgp, skipping the category alcgp=="120+" & agegp=="25-34", using an if else statement.
5. Using a for loop simulate the flip a coin twenty times, keeping track of the individual outcomes (1 = heads, 0 = tails) in a vector that you predefine.



```{r}


```


## apply, lapply, etc...

1. define the following dataset:

```{r}
dataset1 <- cbind(observationA = 16:8, observationB = c(20:19, 6:12))
```

2. Calculate row and column means with apply 
3. Use apply to find the min and the max of each row and column. 
4. Do the same, but use a for cylce. 

```{r}

```


```{r}

```

4. create a list with 3 list elements. Print the length of each element. 

```{r}
lll <- list(list(1,3,5,6,1), list(5,2,1), list(6,2,5,7,3,1,9))
```


6. Multiply each list by 2, find each list's minimum and maximum value. 

```{r}


```

5. Do the same, but use a for cycle on the list. 
6. add names to each listelements. The names should be big, if the element is larger than the medium of the list and small otherwise. The medium should be medium. 


