---
title: "Regular expressions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{REgular expressions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Regular expressions

Regular expressions are a concise and flexible tool for describing patterns in strings. This vignette describes the key features of stringr’s regular expressions, as implemented by stringi. It is not a tutorial, so if you’re unfamiliar regular expressions, I’d recommend starting at https://r4ds.had.co.nz/strings.html. If you want to master the details, I’d recommend reading the classic Mastering Regular Expressions by Jeffrey E. F. Friedl.

Regular expressions are the default pattern engine in stringr. That means when you use a pattern matching function with a bare string, it’s equivalent to wrapping it in a call to regex():

## Basic matches

The simplest patterns match exact strings:

```{r}
x <- c("apple", "banana", "pear", "tapas")
grep( "an", x)
```
```{r}

grep("a.a", x)
grep("a\\.a", x)
```
You can also create your own __character classes__ using `[]`:

* `[abc]`: matches a, b, or c.
* `[a-z]`: matches every character between a and z 
   (in Unicode code point order).
* `[^abc]`: matches anything except a, b, or c.
* `[\^\-]`: matches `^` or `-`.

There are a number of pre-built classes that you can use inside `[]`:

* `[:punct:]`: punctuation.
* `[:alpha:]`: letters.
* `[:lower:]`: lowercase letters.
* `[:upper:]`: upperclass letters.
* `[:digit:]`: digits.
* `[:xdigit:]`: hex digits.
* `[:alnum:]`: letters and numbers.
* `[:cntrl:]`: control characters.
* `[:graph:]`: letters, numbers, and punctuation.
* `[:print:]`: letters, numbers, punctuation, and whitespace.
* `[:space:]`: space characters (basically equivalent to `\s`).
* `[:blank:]`: space and tab.


## Alternation

`|` is the __alternation__ operator, which will pick between one or more possible matches. For example, `abc|def` will match `abc` or `def`:


```{r}
grep( "abc|def", c("abc", "def", "ghi"))
```



* `^` now matches the start of each line. 

* `$` now matches the end of each line.

* `\A` matches the start of the input.

* `\z` matches the end of the input.

* `\Z` matches the end of the input, but before the final line terminator, 
  if it exists.
  
## Repetition

You can control how many times a pattern matches with the repetition operators:

* `?`: 0 or 1.
* `+`: 1 or more.
* `*`: 0 or more.


* `{n}`: exactly n
* `{n,}`: n or more
* `{n,m}`: between n and m

## `grep`, `grepl`, `gsub`


```{r}
gsub("ap", "b", x)
```

