---
title: "Reproducible programming"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducible programming}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Reproducible programming


> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))


## General rules

1. the scripts should be written in a way, that either you, or someone else will be able to run it anytime.

2. Clean structure, according to the best programming practices

* Variable names according the convention (use _. not " " or -, don't use keywords)
* `<-` instead of `=`
* spaces before and after
* Indentation
* curly brackets
* don't store unnecessary objects
* avoid repetition

- comment your codes in detail
- Use sessionInfo() to know your package versions
- Use as generalized code as possible - use names and regular expressions instead of indices. 
- if possible, define your paths at the beginning and use paste() or file.path() later. Or use relative paths
- try to do everything from R, because changes in e.g. excel are undocumented.




## Rmarkdown, R notebooks

* Rmarkdown document, with code chuncks
* Markdown language is easy to learn, see the cheatsheat
* many templates available - flexible themes
* Description, codes and figures at the same place. 
* Figures are saved separately as well - or use links
* Appearance of the code depends on the chunk options.

Show code and output

```{r}

plot(cars)

```

Show only output

```{r echo=FALSE}
plot(cars)
```

Show code only, don't run

```{r eval=FALSE}
plot(cars)
```

* Suppress warnings and messages, if needed.
* Use cheatsheets
* easy to add pictures, links, etc.
* easy to add nice tables

```{r, results='asis'}
library(dplyr)
data("mtcars")
knitr::kable(mtcars, format = "html")
print(xtable::xtable(mtcars), type = 'html')
 mtcars %>%
 	DT::datatable(filter = 'top', 
          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5),  
          rownames = FALSE)
```


```{r}
sessionInfo()
```




